---
title: "marking-mcq-code"
author: "David Reinstein"
date: "11/2/2019"
output: html_document
---

# Outline of code and procedure

1. mark_scheme: ms <- Input 'marking.csv' file of q_label, point_value, d_partial_credit, num_ans (number of possible answers per question, typically 5), c_ans (correct_answers), alt_ans (alternate correct, where non-blank), free_marks (grant everyone this share of marks for a question as a minimum)

- Convert correct_answers, alt_correct to character lists

2. student_responses: sr <- Input (key sheet and rows of) Excel file of student_id, r_q_[label] (responses to all questions, same order as for marking.csv) as produced by Exeter MCQ administrative processing; each row is a student

3. Clean student_responses file to aid processing:

- dataframe or matrix; each r_q[label] is a list of responses chosen ("A","C","E") etc
- clean each response to have the same format as mark_scheme$correct_answers (capital letters etc)

4. Create additional 'sr$m_pct_q_[label]' columns^[note, this could also be done by creating a different matrix, or a multidimensional array] in sr, one for each question, calculating 'share of marks to be awarded' as follows

Comparing `ms$c_ans` and `student_responses$response_q` element-by-element

- If d_partial_credit = 0, then sr$m_pct_q_[label] <= ifelse 'exact match' 1,0

- If d_partial_credit = 1, then 
  - Share of correct: (sum 'list elements in common') / length(`ms$q_label`)
  - less 'Share of incorrect': ((length(`ms$q_label`) - sum 'list elements in common')) / `(ms$num_ans - length(ms$c_ans)`)

- if `length$alt_ans>0` do the previous step again wrt `alt_ans`, assigning a column sr$alt_m_pct_q_[label]

- set `sr$max_m_pct_q_[label]` to max of 0, `sr$alt_m_pct_q_[]`,  `sr$m_pct_q_[]`, `sr$free_marks` (note no negative marks for a question, award best possible mark for all alternate answers)

5. Take product of `ms$point_value`  and `sr$max_m_pct_q_[label]` element by element for each student and question, creating `sr$mark_q_[label]` columns.

6. `sr$total_mark <- ` Sum `sr$mark_q_[label]` for each student to get overall mark.

7. Output as organised excel sheet (and other formats to give individual students their breakdowns?)


```{r local_values}

folder <- "/Users/yosemite/OneDrive - University of Exeter/exeter_teaching/marking/beem101_midterm_marking"

markscheme <- "mark_scheme.csv"
student_responses <- "BEEM101 Data.xlsx"
adhoc_adjust <- 0

outfile <-  "beem101_midterm_results.csv"

maxmark <- 50
```


```{r setup, include=TRUE, warning=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
require("knitr")

opts_knit$set(root.dir = folder)

library(pacman)
#knitr,dplyr,tidyverse,labelled,citr,reporttools,magrittr,glue,huxtable,experiment,dataMaid,broom,janitor,here,xRStata,estimatr,xtable
p_load(dplyr,magrittr,purrr,tidyverse,tidyr,broom,janitor, here,glue,dataMaid,readr,lubridate,summarytools, httr,jsonlite,rlist,XML, readxl)

```


Just go to the relevant directory 
```{r}

setwd(folder)

```


# 1. mark_scheme: ms <- Input 'marking.csv' file

```{r input-mark-scheme}

ms <- read.csv(markscheme, header=TRUE,  stringsAsFactors=FALSE, row.names = 1) 

```

Convert correct_answers, alt_correct to character lists:

```{r}

ms$c_ans <- lapply(strsplit(as.character(ms$c_ans),split=','),trimws)
ms$alt_ans <- lapply(strsplit(as.character(ms$alt_ans),split=','),trimws)

```

Count number of correct answers
```{r}

for (i in 1:NROW(ms)) {          
ms$numc[i] <- length(ms$c_ans[[i]])
ms$numc_a[i] <- length(ms$alt_ans[[i]])
}
```


# 2. input student_responses
3. Clean student_responses file

sr <- input mark sheet, remove parens

```{r}

sr <- read_excel(student_responses) 

sr <- lapply(sr, gsub, pattern = ")", replacement = "", fixed = TRUE)
sr <- lapply(sr, gsub, pattern = "(", replacement = "", fixed = TRUE)
sr <- lapply(sr, gsub, pattern = "BLANK", replacement = "", fixed = TRUE)

sr <- as_tibble(sr)
```

Convert all but the first column to a list column:
```{r}
makeitlist <- function(x) (lapply(strsplit(as.character(x),split=','),trimws))

for (i in 2:length(sr)) {          
  sr[[i]] <- makeitlist(sr[[i]]) 
}

```

# Marking process

4. Create additional 'sr$m_pct_q_[label]' 'share of marks to be awarded' as follows

Compare `ms$c_ans` and `sr$response_q` element-by-element and count number of matches


```{r count-matches-to-correct-ans}

#This seems to work, element by element...
#length(intersect(sr$questions3[[2]], ms[[c("questions3"),"c_ans"]]))
#length(intersect(sr$questions3[[1]], ms[[c("questions3"),"c_ans"]])) 

#now iterate it over all questions and over all students

for(q in rownames(ms)) {   #iterate over all questions 
  for (i in 1:NROW(sr)) {  #over all students
    sr[[glue("cr_ans_",q)]][[i]] <- length(intersect(sr[[q]][[i]], ms[[c(q),"c_ans"]])) 
  }
}
```




- If d_partial_credit = 0, then sr$m_pct_q_[label] <= ifelse 'exact match' 1,0

- If d_partial_credit = 1, then 
  - Share of correct: (sum 'list elements in common') / length(`ms$q_label`)
  - less 'Share of incorrect': ((length(`ms$q_label`) - sum 'list elements in common')) / `(ms$num_ans - length(ms$c_ans)`)

```{r award-marks}

#- If d_partial_credit = 0, then sr$m_pct_q_[label] <= ifelse 'exact match' 1,0

for(q in rownames(ms)) {   #iterate over all questions 
  for (i in 1:NROW(sr)) {  #over all students
    sr[[glue("sh_cor",q)]][[i]]   <- sr[[glue("cr_ans_",q)]][[i]]/length(ms[[c(q),"c_ans"]]) #share of correct answers attained
    sr[[glue("sh_incor",q)]][[i]] <-  (length(sr[[q]][[i]])-sr[[glue("cr_ans_",q)]][[i]])/(ms[[c(q),"num_choices"]] - length(ms[[c(q),"c_ans"]])) #share of incorrect answers chosen
    sr[[glue("net_cor",q)]][[i]] <- max(0,sr[[glue("sh_cor",q)]][[i]]-sr[[glue("sh_incor",q)]][[i]])  #net positive, minimum 0
    sr[[glue("mk_",q)]][[i]] <- sr[[glue("net_cor",q)]][[i]] *ms[[c(q),"point_value"]] #multiply by point value
    sr[[glue("mk_",q)]][[i]] <- sr[[glue("mk_",q)]][[i]]*(ms[[c(q),"d_partial_credit"]]) +  sr[[glue("mk_",q)]][[i]]*(1-(ms[[c(q),"d_partial_credit"]]))*(sr[[glue("sh_incor",q)]][[i]]==0) # Set imperfect answers to 0 where no partial credit is granted 
  }
}

sr <- sr %>%
  select(-matches("sh_cor|sh_incor|net_|cr_"))

```

## Marks based on alternate correct answers:
```{r, include=FALSE}

for(q in rownames(ms)) {   #iterate over all questions 
  for (i in 1:NROW(sr)) {  #over all students
    sr[[glue("cr_ans_alt",q)]][[i]] <- length(intersect(sr[[q]][[i]], ms[[c(q),"alt_ans"]])) 
  }
}

for(q in rownames(ms)) {   #iterate over all questions 
  for (i in 1:NROW(sr)) {  #over all students
    sr[[glue("sh_cor_alt",q)]][[i]]   <- sr[[glue("cr_ans_alt",q)]][[i]]/length(ms[[c(q),"alt_ans"]]) #share of correct answers attained
    sr[[glue("sh_incor_alt",q)]][[i]] <-  (length(sr[[q]][[i]])-sr[[glue("cr_ans_alt",q)]][[i]])/(ms[[c(q),"num_choices"]] - length(ms[[c(q),"alt_ans"]])) #share of incorrect answers chosen
    sr[[glue("net_cor_alt",q)]][[i]] <- max(0,sr[[glue("sh_cor_alt",q)]][[i]]-sr[[glue("sh_incor_alt",q)]][[i]])  #net positive, minimum 0
    sr[[glue("mk_alt",q)]][[i]] <- sr[[glue("net_cor_alt",q)]][[i]] *ms[[c(q),"point_value"]] #multiply by point value
  }
}

```

## Take max of marking versions:

```{r,}

for(q in rownames(ms)) {   #iterate over all questions 
  for (i in 1:NROW(sr)) {  #over all students
    sr[[glue("mk_max",q)]][[i]] <- max(sr[[glue("mk_",q)]][[i]],   sr[[glue("mk_alt",q)]][[i]],ms[[c(q),"free_marks"]],na.rm = TRUE)
  }
  }

```

Ad-hoc sympathy points ... if you chose C-D for questions6 you gain 1.5 marks
```{r}
  for (i in 1:NROW(sr)) {  #over all students
  sr$q6_cd[i] <- length(intersect(sr$questions6[[i]],c("C","D")))==2
  sr$mk_maxquestions6[i]  <- min(3, sr$mk_maxquestions6[i]+1.5*sr$q6_cd[i])
  }

```


## Generate total marks
```{r}
sr <- sr %>%
  mutate(
    total = rowSums(.[grep("mk_max*", names(.))], na.rm = TRUE),
    total_adjust = rowSums(.[grep("mk_max*", names(.))], na.rm = TRUE)+ adhoc_adjust,
        tot_adj_share  = total_adjust/maxmark
  )
```

## Merge long-answer marks
```{r}

long_ans <- read.csv("101_midterm_19-Grid view_done.csv", header=TRUE,  stringsAsFactors=FALSE)  %>% as_tibble()

sr_all <- sr %>% mutate(Cand_num = as.numeric(`Student Candidate Number`)) %>%
  mutate(Cand_num=recode(Cand_num, 
                         `51066`=151066)) %>%
  full_join(long_ans, ., by=c("Cand_num")) %>% 
  select(Cand_num, matches("total_adjust|mk_max|Mark|total.for"),starts_with("q"),-starts_with("qu")) %>%
  mutate(total_mark=total.for.part.B..formula.+total_adjust) %>%
  select(Cand_num,total_mark,starts_with("total"), everything()) %T>%
  print()


```

# Flag scripts with missing or strange entries:
```{r}
srq <- sr %>%  select(1,starts_with("questions"))  
srq$q_missing <- 0
  
for(q in colnames(srq[2:NCOL(srq)-1])) {   #iterate over all questions 
   for (i in 1:NROW(srq)) {  #over all students
      srq[[glue("l_",q)]][[i]]   <-  (length(srq[[glue(q)]][[i]])==0)        
   }
}

srq <- srq %>% 
  mutate(
    q_missing =    rowSums(select(., starts_with("l_")))
  )
     
srq %>% filter(q_missing>0) %>% select(starts_with("Student")) %>%  as.matrix() %>% sort() %>%  print() 
  
#filter((length(questions2[[]]) == 0)(questions2)) #if any answer has a blank or a character other than A-E

```


# Exporting and reporting
## Export spreadsheet of marks

```{r}

sr_all %>% 
  rename_all(
    list(
    ~stringr::str_replace_all(.,c('maxquestions'='q','adjust'='part_a','.for.part.B..formula.'='_part_B','Mark.1...out.of.'='mark_B_first_q_from_', 'Mark.2...out.of.'='mark_B_second_q_from_','q(.)c'='q\\1_comments')) 
    )
  ) %>%
  select(-"q6_cd") %>%
  arrange(Cand_num) %>%
write_excel_csv(file.path(folder,outfile)) 

```

## Summary statistics

```{r}

options(digits = 4)

sr_all %>% 
  dplyr::summarise(mean=mean(total_mark, na.rm = TRUE),
                   median=median(total_mark, na.rm = TRUE),
                   p25 = quantile(total_mark,.25, na.rm = TRUE),
                   p75 = quantile(total_mark,.75, na.rm = TRUE)) 

#marks by question
sr %>% select(starts_with("mk_maxq")) %>% 
  dplyr::summarise_all(mean) %>% transpose()
```


## Export one file per student-row (feedback)
```{r}

sr_all %>% 
  rename_all(
    list(
    ~stringr::str_replace_all(.,c('maxquestions'='q','adjust'='part_a','.for.part.B..formula.'='_part_B','Mark.1...out.of.'='mark_B_first_q_from_', 'Mark.2...out.of.'='mark_B_second_q_from_','q(.)c'='q\\1_comments')) 
    )
  ) %>%
  dplyr::select(-q6_cd) %>%
  dplyr::arrange(Cand_num) %>%
  mutate(index = 1:n()) %>%
  rowwise %>%
  do(write.table(., 
                 file=paste0("candidate_", .$Cand_num, ".tsv"),
                 row.names=FALSE, sep="\t", eol = "\r\n" ) ) 

for (i in 1:nrow(sr_all)) {
    sink(paste0(folder, "/feedback_cand_", sr_all$Cand_num[i], ".txt"))
  print(i)
      print(sr_all[glue('{i}'),])
    sink()
}

for (i in seq_along(x)) {
  name <- names(x)[[i]]
  value <- x[[i]]
}


library(dplyr)
mtcars %>%
  mutate(index = 1:n() ) %>% 

  
  do(write.table(., 
                 file=paste0("matchers", .$index, ".txt"),
                 row.names=FALSE, sep="\t") )

mtcars %>%
  mutate(index = 1:n() ) %>%
  rowwise %>%
  do(write.table(., 
                 file=paste0("matchers", .$index, ".txt"),
                 row.names=FALSE, sep="\t") )
  
```

## Input random ordering from uni, and put the marks in this order, and then export again

```{r}
ordering <- read_excel("cand_number_ordering.xlsx", col_names = FALSE) %>% 
lapply(., gsub, pattern = "#", replacement = "", fixed = TRUE) %>% as.data.frame()

ordering$id  <- 1:nrow(ordering)

test <- sr_all %>% select(1,"total_mark") %>%  merge(., ordering,  by.x="Cand_num", by.y="...1",   all = TRUE)  %>% 
  arrange(id) %>%
  write_excel_csv(file.path(folder,"ordered_marks.csv")) 


```




