---
title: "marking-mcq-code"
author: "David Reinstein"
date: "11/2/2019"
output: html_document
---

# Outline of code and procedure

1. mark_scheme: ms <- Input 'marking.csv' file of q_label, point_value, d_partial_credit, num_ans (number of possible answers per question, typically 5), c_ans (correct_answers), alt_ans (alternate correct, where non-blank), free_marks (grant everyone this share of marks for a question as a minimum)

- Convert correct_answers, alt_correct to character lists

2. student_responses: sr <- Input (key sheet and rows of) Excel file of student_id, r_q_[label] (responses to all questions, same order as for marking.csv) as produced by Exeter MCQ administrative processing; each row is a student

3. Clean student_responses file to aid processing:

- dataframe or matrix; each r_q[label] is a list of responses chosen ("A","C","E") etc
- clean each response to have the same format as mark_scheme$correct_answers (capital letters etc)

4. Create additional 'sr$m_pct_q_[label]' columns^[note, this could also be done by creating a different matrix, or a multidimensional array] in sr, one for each question, calculating 'share of marks to be awarded' as follows

Comparing `ms$c_ans` and `student_responses$response_q` element-by-element

- If d_partial_credit = 0, then sr$m_pct_q_[label] <= ifelse 'exact match' 1,0

- If d_partial_credit = 1, then 
  - Share of correct: (sum 'list elements in common') / length(`ms$q_label`)
  - less 'Share of incorrect': ((length(`ms$q_label`) - sum 'list elements in common')) / `(ms$num_ans - length(ms$c_ans)`)

- if `length$alt_ans>0` do the previous step again wrt `alt_ans`, assigning a column sr$alt_m_pct_q_[label]

- set `sr$max_m_pct_q_[label]` to max of 0, `sr$alt_m_pct_q_[]`,  `sr$m_pct_q_[]`, `sr$free_marks` (note no negative marks for a question, award best possible mark for all alternate answers)

5. Take product of `ms$point_value`  and `sr$max_m_pct_q_[label]` element by element for each student and question, creating `sr$mark_q_[label]` columns.

6. `sr$total_mark <- ` Sum `sr$mark_q_[label]` for each student to get overall mark.

7. Output as organised excel sheet (and other formats to give individual students their breakdowns?)



```{r local_values}

folder <- "/Users/yosemite/OneDrive - University of Exeter/exeter_teaching/marking/bee2038_midterm_marking"

markscheme <- "mark_scheme.csv"
student_responses <- "BEE2038 - data - not marked.xlsx"
```



```{r setup, include=TRUE, warning=FALSE, echo=FALSE}

knitr::opts_chunk$set(echo = TRUE)
require("knitr")

opts_knit$set(root.dir = folder)

library(pacman)
#knitr,dplyr,tidyverse,labelled,citr,reporttools,magrittr,glue,huxtable,experiment,dataMaid,broom,janitor,here,xRStata,estimatr,xtable
p_load(dplyr,magrittr,purrr,tidyverse,tidyr,broom,janitor, here,glue,dataMaid,readr,lubridate,summarytools, httr,jsonlite,rlist,XML, readxl)


```


Just go to the relevant directory 
```{r}

setwd(folder)

```


# 1. mark_scheme: ms <- Input 'marking.csv' file

```{r input-mark-scheme}


ms <- read.csv(markscheme, header=TRUE,  stringsAsFactors=FALSE, row.names = 1) %>%
  as_tibble() 

```

- Convert correct_answers, alt_correct to character lists

```{r}

ms$c_ans <- lapply(strsplit(as.character(ms$c_ans),split=','),trimws)
ms$alt_ans <- lapply(strsplit(as.character(ms$alt_ans),split=','),trimws)

```


2. student_responses: sr <- input mark sheet, remove parens, make list columns

```{r}

sr <- read_excel(student_responses) 

sr <- lapply(sr, gsub, pattern = ")", replacement = "", fixed = TRUE)
sr <- lapply(sr, gsub, pattern = "(", replacement = "", fixed = TRUE)

sr <- as_tibble(sr)

sr$questions1 <- lapply(strsplit(as.character(sr$questions1),split=','),trimws)
#apply this to all elements but the first one



```

